\textsf{Are large language models superhuman chemists?}

### 2024-10-24 src/tex/ms.tex
address humans as volunteers

We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists in our study on average.
However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.

@@

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human) chemists.

@@

To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human) chemists of different specializations. In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework. To provide a baseline, we built a web application that we used to survey experts in chemistry.
    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
    }
    \label{fig:overview_figure}
\end{figure}

@@

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the human experts in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
We then computed the percentage of questions the models or humans answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). Thus, a larger colored area indicates a better performance.

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about \gls{ghs} pictograms Claude 3 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

\section{Conclusions}
On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics.
On the other hand, there are still striking limitations.
For very relevant topics the answers models provide are wrong.
On top of that, many models are not able to reliably estimate their own limitations.
Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the questions we use to evaluate models---and chemists---than about the models themselves.
For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans (when trained on the right training corpus).

>>>>

We curated more than 2,800 question-answer pairs, evaluated leading open- and closed-source LLMs, and found that the best models outperformed the best human chemists (volunteers) in our study on average.
However, the models struggle with some tasks that are easy for human experts and provide overconfident predictions.

@@

However, the rapid increase in capabilities of chemical \gls{ml} models led (even before the recent interest in \glspl{llm}) to concerns about the potential for dual use of these technologies, e.g., for the design of chemical weapons.\autocite{gopal2023releasing, ganguli2022red, Urbina_2022, campbell2023censoring, moulange2023towards, urbina2022teachable}
To some extent, this is not surprising as any technology that, for instance, is used to design non-toxic molecules can also be used inversely to predict toxic ones (even though the synthesis would still require access to controlled physical resources and facilities).
Still, it is essential to realize that such models' user base is wider than that of chemistry and materials science experts who can critically reflect on every output such models produce.
For example, many students frequently consult these tools---perhaps even to prepare chemical experiments.\autocite{Intelligent.com_2023}
This also applies to users from the general public, who might consider using \glspl{llm} to answer questions about the safety of chemicals.
Thus, for some users, misleading information---especially about safety-related aspects---might lead to harmful outcomes.
However, even for experts, chemical knowledge and reasoning capabilities are essential as they will determine the capabilities and limitations of their models in their work, e.g., in copilot systems for chemists.
Unfortunately, apart from exploratory reports such as by prompting leading models with various scientific questions,\autocite{ai4science2023impact} there is little systematic evidence on how \glspl{llm} perform compared to expert (human volunteers) chemists.

@@

To contextualize the scores, we also surveyed \variable{output/number_experts.txt} experts in chemistry on a subset of the benchmark corpus to be able to compare the performance of current frontier models with (human volunteers) chemists of different specializations. In parts of the survey, the participants were also allowed to use tools such as web search to create a realistic setting.
Our results indicate that current frontier models perform \enquote{superhuman} on some aspects of chemistry but, in many cases, including safety-related ones, might be very misleading.
We find that there are still limitations for current models to overcome, such that they can be directly applied in autonomous systems for chemists.

@@

\begin{figure}
    \includegraphics[width=\textwidth]{figures/overview_figure.pdf}
    \caption{\textbf{Overview of the \chembench framework.} The figure shows the different components of the \chembench framework.
    The framework's foundation is the benchmark corpus that consists of many questions and answers that we manually or semi-automatically compiled from various sources.
    Questions are classified based on topics, required skills (reasoning, calculation, knowledge, intuition), and difficulty levels.
    We then used this corpus to evaluate the performance of various models and tool-augmented systems using a custom framework. To provide a baseline, we built a web application that we used to survey experts (human volunteers) in chemistry.
    The results of the evaluations are then compiled in publicly accessible leaderboards, which we propose as a foundation for evaluating future models.
    }
    \label{fig:overview_figure}
\end{figure}

@@

\paragraph{Overall system performance}
To understand the current capabilities of \glspl{llm} in the chemical sciences, we evaluated a wide range of leading models\autocite{Huggingface} on the \chembench corpus, including systems augmented with external tools.
An overview of the results of this evaluation is shown in \Cref{fig:human_vs_models_bar} (all results can be found in \Cref{tab:performance_table} and \Cref{tab:performance_table_human_subset}).
In this figure, we show the percentage of questions that the models answered correctly.
Moreover, we show the worst, best, and average performance of the experts (human volunteers) in our study, which we obtained via a custom web application (\url{chembench.org}) that we used to survey the experts.
Remarkably, the figure shows that the leading \gls{llm}, \oone, outperforms the best human in our study in this overall metric by almost a factor of two.
Many other models also outperform the average human performance.
Interestingly, \LlamaThreeOneFourZeroFiveBInstruct shows performance that is close the leading proprietary models, indicating that new open source models can be competitive with the best proprietary models also in chemical settings.

@@

\paragraph{Performance per topic} To obtain a more detailed understanding of the performance of the models, we also analyzed the performance of the models in different subfields of the chemical sciences.
For this analysis, we defined a set of topics (see \Cref{sec:meth-topic}) and  classified all questions in the \chembench corpus into these topics.
We then computed the percentage of questions the models or experts (human volunteers) answered correctly for each topic and show them in \Cref{fig:all_questions_models_completely_correct_radar_human}.
In this spider chart, the worst score for every dimension is zero (no question answered correctly), and the best score is one (all questions answered correctly). Thus, a larger colored area indicates a better performance.

@@

 In \Cref{fig:confidence_vs_performance}, we show that for some models, there is no significant correlation between the estimated difficulty and whether the models answered the question correctly or not.
 For applications in which humans might rely on the models to provide answers with trustworthy uncertainty estimates, this is a concerning observation highlighting the need for critical reasoning in the interpretation of the model's outputs.\autocite{Li_2023, miret2024llms}
 For example, for the questions about the safety profile of compounds, \GPTFour reported an average confidence of \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_correct_overall.txt} (on a scale of 1--5) for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_correct_overall.txt} questions it answered correctly and \variable{output/model_confidence_performance/gpt-4_is_pictograms_average_confidence_incorrect_overall.txt} for the \variable{output/model_confidence_performance/gpt-4_is_pictograms_num_incorrect_overall.txt} questions it answered incorrectly.
 While, on average, the verbalized confidence estimates from Claude 3.5 seem better calibrated (\Cref{fig:confidence_vs_performance}), they are still misleading in some cases.
 For example, for the questions about \gls{ghs} pictograms Claude 3.5 returns an average score of \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_correct_overall.txt} for correct answers and \variable{output/model_confidence_performance/claude3_is_pictograms_average_confidence_incorrect_overall.txt} for incorrect answers.

\section{Conclusions}
On the one hand, our findings underline the impressive capabilities of \glspl{llm} in the chemical sciences: Leading models outperform domain experts in specific chemistry questions on many topics.
On the other hand, there are still striking limitations.
For very relevant topics, the answers that models provide are wrong.
On top of that, many models are not able to reliably estimate their own limitations.
Yet, the success of the models in our evaluations perhaps also reveals more about the limitations of the questions we use to evaluate models---and chemists---than about the models themselves.
For instance, while models perform well on many textbook questions, they struggle with questions that require some more reasoning about chemical structures (e.g., number of isomers or \gls{nmr} peaks).
Given that the models outperformed the average human in our study, we need to rethink how we teach and examine chemistry.
Critical reasoning is increasingly essential, and rote solving of problems or memorization of facts is a domain in which \glspl{llm} will continue to outperform humans (when trained on the right training corpus).
